Covariate analysis of MSBB microarray data

### 2016-01-04 code/Rmd/MSSM_Array.Rmd
Minor modification in MSSB Array covariates adjustments

```{r covariate analysis}
# Function to do covariates adjustments
covariatesAdjustments <- function(exprId, covId){
  # Download expression data
  expr <- read.table(synGet(exprId)@filePath, header=T, sep='\t', check.names = F)
  expr <- expr[expr$Gene.Symbol!='',]
  rownames(expr) <- expr$ID
  ANNOTATIONS <- dplyr::select(expr,ID,GB_ACC,Gene.Symbol,ENTREZ_GENE_ID)
  expr <- dplyr::select(expr,-c(ID,GB_ACC,Gene.Symbol,ENTREZ_GENE_ID))
  rownames(expr) = ANNOTATIONS$ID

  Collapsed.obj <- collapseRows(expr,ANNOTATIONS$Gene.Symbol,ANNOTATIONS$ID)
  expr <- Collapsed.obj$datETcollapsed

@@

  # Plot significant data
  print(preAdjustedSigCovars$PC_res[[2]]$plotData)
  print(paste('Significant covariates are:',paste(preAdjustedSigCovars$significantCovars,collapse=',')))

  # Get design matrix
  DESIGN <- getDesignMatrix(covar[,'Dx',drop=F], Intercept = F)
  print(paste('Fitting linear model with the following coefficients:',paste(colnames(DESIGN$design),collapse=',')))

@@

  if (length(postAdjustedSigCovars$significantCovars) != 0)
    print(paste('Following coefficients have to be included in the model:',paste(postAdjustedSigCovars$significantCovars,collapse=',')))

  return(list(EXPR = expr,
              COVAR = covar,
              FIT = fit, 
              RESIDUAL = residualFit, 
              DESIGN = DESIGN, 
              ANNOTATIONS = ANNOTATIONS))
}
```
### Middle Temporal Analysis
```{r middle.temporal, cache=TRUE, fig.height=10, fig.width=15}
# SynIDs of expression and covariates matrix
EXP_ID <- c('Middle Temporal' = 'syn3191101',
            'Superior Temporal' = 'syn3191103',
            'Inferior Temporal' = 'syn3191099',
            'Superior Parietal Lobule' = 'syn3191119',
            'Caudal Anterior Cingulate' = 'syn3191107',
            'Parahippocampal Gyrus' = 'syn3191109',
            'Frontal Pole' = 'syn3191107')
COV_ID <- 'syn3205399'

# Obtain adjusted expession
adjustedExpression <- lapply(EXP_ID, covariatesAdjustments, COV_ID)
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE}
# Code
CODE <- Folder(name = "MSBB Microarray", parentId ="syn5570248")
CODE <- synStore(CODE)

>>>>

```{r fxns}
# Function to do covariates adjustments
covariatesAdjustments <- function(exprId, regionName, covId){
  inline_expr(paste0('Working on ',regionName), "md")

  # Download expression data
  expr <- read.table(synGet(exprId)@filePath, header=T, sep='\t', check.names = F)
  expr <- expr[expr$Gene.Symbol!='',]
  rownames(expr) <- expr$ID
  ANNOTATIONS <- dplyr::select(expr,ID,GB_ACC,Gene.Symbol,ENTREZ_GENE_ID)
  expr <- dplyr::select(expr,-c(ID,GB_ACC,Gene.Symbol,ENTREZ_GENE_ID))
  rownames(expr) = ANNOTATIONS$ID

  # Collapse probe ids to hgnc symbols
  Collapsed.obj <- collapseRows(expr,ANNOTATIONS$Gene.Symbol,ANNOTATIONS$ID)
  expr <- Collapsed.obj$datETcollapsed

@@

  # Plot significant data
  print(preAdjustedSigCovars$PC_res[[2]]$plotData)
  inline_expr(paste('Significant covariates are:',paste(preAdjustedSigCovars$significantCovars,collapse=',')), "md")

  # Get design matrix
  DESIGN <- getDesignMatrix(covar[,'Dx',drop=F], Intercept = F)
  inline_expr(paste('Fitting linear model with the following coefficients:',paste(colnames(DESIGN$design),collapse=',')), "md")

@@

  if (length(postAdjustedSigCovars$significantCovars) != 0)
    inline_expr(paste('Following coefficients have to be included in the model:',paste(postAdjustedSigCovars$significantCovars,collapse=',')), "md")

  return(list(EXPR = expr,
              COVAR = covar,
              FIT = fit, 
              RESIDUAL = residualFit, 
              DESIGN = DESIGN, 
              ANNOTATIONS = ANNOTATIONS))
}
```
### Extract expression data and adjust covariates 
```{r covariate.analysis, cache=TRUE, fig.height=10, fig.width=15}
# SynIDs of expression and covariates matrix
EXP_ID <- c('Middle Temporal' = 'syn3191101',
            'Superior Temporal' = 'syn3191103',
            'Inferior Temporal' = 'syn3191099',
            'Superior Parietal Lobule' = 'syn3191119',
            'Caudal Anterior Cingulate' = 'syn3191107',
            'Parahippocampal Gyrus' = 'syn3191109',
            'Frontal Pole' = 'syn3191107')
COV_ID <- 'syn3205399'

# Obtain adjusted expession
adjustedExpression <- mapply(covariatesAdjustments, EXP_ID, names(EXP_ID), MoreArgs= list(COV_ID), SIMPLIFY=T)
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE}
# Code
CODE <- Folder(name = "MSBB Microarray", parentId ="syn5570248")
CODE <- synStore(CODE)
